\section{Introduction}
Modern cosmology rests on two pillars: General Relativity and Quantum Field Theory. Yet these pillars are at war. The ``Vacuum Catastrophe"—the $10^{120}$ discrepancy between the predicted quantum vacuum energy and the observed cosmological constant—is the most famous failure. More recently, high-precision measurements have revealed significant tensions within the standard $\Lambda$CDM model itself. The ``$H_0$ Tension" shows a $5\sigma$ conflict between early-universe and late-universe measurements of the expansion rate, while the ``$S_8$ Tension" reveals a related discrepancy in the observed clumpiness of cosmic structure.

These are not minor fitting issues; they are foundational cracks suggesting that our model of the vacuum is incomplete. In this paper, we propose that these tensions are not evidence for new particles or modified gravity, but are the predictable consequences of treating the universe as a finite-capacity, information-processing lattice.

Building on the $E_8$-Persistence framework established in Papers I-III, this work presents a complete cosmological model from the Big Bang to black holes. We will demonstrate that:
\begin{enumerate}
    \item \textbf{The parameters of inflation} are fixed by the lattice's initialization sequence, not an arbitrary inflaton potential.
    \item \textbf{The $H_0$ and $S_8$ tensions} are dual manifestations of a single, geometrically-derived \textbf{Lattice Load Factor ($\mathcal{L}$)}, representing the systemic latency of a matter-filled vacuum.
    \item \textbf{The laws of black hole thermodynamics} emerge from the same geometric invariants that define the Standard Model, resolving the Information Paradox.
    \item \textbf{The ``Dark Sector"} is not composed of new particles, but is the geometric mass (Dark Matter) and entropic exhaust (Dark Energy) of the vacuum substrate itself.
\end{enumerate}
This paper unifies the physics of the primordial and modern universe under the single principle of information conservation on a discrete substrate, replacing the fitted parameters of $\Lambda$CDM with parameter-free geometric derivations.