\documentclass[aps,prd,twocolumn,showpacs,superscriptaddress,groupedaddress,nofootinbib,english,floatfix]{revtex4-2}  

% Packages you will need
\usepackage{amsmath}  % Math formulas
\usepackage{amssymb}  % Math symbols (like \mathbb, \mathcal)
\usepackage{graphicx} % Images
\usepackage[pdfencoding=auto]{hyperref} % Hyperlinks
\usepackage{bm}       % Bold math
\usepackage{booktabs} % For professional table formatting
\usepackage{array}    % For better column definitions
\usepackage{tabularx, siunitx}
\usepackage[ngerman, main=english]{babel}
\usepackage{catchfilebetweentags} 
\usepackage{siunitx}
\sisetup{
    separate-uncertainty = true,  % Forces the ± symbol
    multi-part-units = single,    % Wraps units: (1.2 ± 0.1) kg
    sticky-per = true             % Helps with complex units
}
\usepackage{comment}
\usepackage[noabbrev]{cleveref} % Load this LAST in your preamble

\makeatletter
\let\l@de\l@ngerman
\makeatother
\makeatletter
\let\l@en\l@english
\makeatother

\begin{document}

% Title Area
\title{The \texorpdfstring{$E_8$}{E8}-Persistence Theory IV: Informational Cosmology and Lattice Limits}


\author{Kate Lenore Meyer}
\affiliation{Independent Researcher}
\email{kate.physics@meyerhome.net}

\date{\today}

\input{paper/0_abstract}

\maketitle % Generates the title block

% Main Content
\input{paper/1_introduction}

% Main Content
\input{paper/2_init}
\input{paper/3_tensions}
\input{paper/4_blackholes}
\input{paper/5_dark_sector}

\input{paper/audit}
\input{paper/conclusion}

%\appendix
%\input{paper/12_geometric_audits}

% Bibliography
\bibliography{paper/references}

\begin{thebibliography}{9}
    \bibitem{Planck2018}
    Planck Collaboration, "Planck 2018 results. VI. Cosmological parameters," \textit{Astron. Astrophys.} 641, A6 (2020).
    
    \bibitem{Riess2022}
    A. G. Riess et al. (SH0ES Collaboration), "A Comprehensive Measurement of the Local Value of the Hubble Constant with 1 km/s/Mpc Uncertainty from the Hubble Space Telescope and the SH0ES Team," \textit{Astrophys. J. Lett.} 934, L7 (2022).
    
    \bibitem{DES-Y3}
    T. M. C. Abbott et al. (DES Collaboration), "Dark Energy Survey Year 3 Results: Cosmological Constraints from Galaxy Clustering and Weak Lensing," \textit{Phys. Rev. D} 105, 023520 (2022).
\end{thebibliography}


\end{document}